<center>
# Fall 2024
## MATH 261: Intro to Data Science
### Assignment #2 | Total points: 10

### <span style="color:limegreen">**Type your name here **</span>
<span style="color:red">**Date:** Wednesday, 9/4/2024</span>
</center>

\

<h3><span style="color:red">**Important Instructions:**</span></h3>
- **It is mandatory to consistently uphold the DePauw integrity policy,** <span style="text-decoration:underline"> 
[DePauw's Academic Integrity Policy.](https://www.depauw.edu/academics/academic-resources/academic-integrity/)</span>


-	**Refrain from utilizing resources like ChatGPT or any other materials that may lead to plagiarism concerns.**
-	**Carry out your tasks autonomously and avoid sharing any documents via email or other communication platforms.**
-	**Please be aware that appropriate actions will be implemented in case any violation is identified.**

- **Show both your R codes and results in the HTML file.**\
- **Save your .RMD file/.HTML file as your Last Name_A2.RMD/Last Name_A1.HTML**\
- **Submit your Assignment #2 (both .RMD and .HTML files) on Moodle through the appropriatel link.** \

___

\
\


<span style="color:blue">**Question #1:**</span>
\
Type the following:
<div style="float: right">[4 pts.]</div>
\
\
<h4>**Standard Score (Z-Score):**</h4>

A **Z-score** (also known as a standard score) is a statistical measurement that describes a value’s relationship to the mean of a group of values. It represents the number of standard deviations a data point is from the mean.

The formula to calculate the Z-score of a value \( x \) is:

\[
Z = \frac{x - \mu}{\sigma}
\]

Where:

- \( x \) is the data point.
\
- \( \mu \) is the mean of the data set.
\
- \( \sigma \) is the standard deviation of the data set.
\

A Z-score can be positive or negative:

- A **positive Z-score** indicates that the data point is above the mean.
\
- A **negative Z-score** indicates that the data point is below the mean.
\
- A **Z-score of 0** indicates that the data point is exactly at the mean.

\
Z-scores are commonly used in statistics to understand how far away a specific value is from the mean, and they are also useful for comparing data points from different distributions.


\

<span style="color:blue">**Question #2:**</span>
\
Suppose you work at a bank as a Data Scientist and your task is to automate the loan approval process based on applicant credit scores. The credit scores for 100 applicants are given below:

703, 673, 587, 728, 759, 650, 835, 740, 501, 689, 714,
573, 709, 651, 570, 810, 587, 575, 810, 714, 438, 705,
760, 650, 725, 564, 808, 615, 628, 551, 769, 594, 590,
596, 684, 801, 693, 598, 777, 757, 767, 601, 699, 640,
808, 777, 725, 691, 562, 598, 756, 816, 802, 704, 725, 
807, 567, 775, 603, 560, 558, 754, 560, 648, 584, 738, 
816, 621, 665, 780, 569, 734, 662, 629, 559, 563, 562, 
575, 555, 673, 742, 784, 587, 682, 626, 765, 721, 593, 
730, 811, 632, 622, 778, 665, 597, 615, 742, 729, 609,
782

\

Categorize the credit scores as follows using `if` statements and finally count the credit scores within each category.
<div style="float: right">[6 pts.]</div>

\

`Very poor` if the score between 300 and 600\
`Poor` if the score between 601 and 630\
`Fair` if the score between 661 and 720\
`Good` if the score between 721 and 780\
`Excellent` if the score between 781 and 850

\
\

```{r}
credit_scores <- c(703, 673, 587, 728, 759, 650, 835, 740, 501, 689, 714, 573, 709, 
                   651, 570, 810, 587, 575, 810, 714, 438, 705, 760, 650, 725, 564, 
                   808, 615, 628, 551, 769, 594, 590, 596, 684, 801, 693, 598, 777, 
                   757, 767, 601, 699, 640, 808, 777, 725, 691, 562, 598, 756, 816, 
                   802, 704, 725, 807, 567, 775, 603, 560, 558, 754, 560, 648, 584, 
                   738, 816, 621, 665, 780, 569, 734, 662, 629, 559, 563, 562, 575, 
                   555, 673, 742, 784, 587, 682, 626, 765, 721, 593, 730, 811, 632, 
                   622, 778, 665, 597, 615, 742, 729, 609, 782)
```
```{r}
credit_categories <- ifelse(credit_scores >= 300 & credit_scores <= 600, "Very Poor",
                     ifelse(credit_scores >= 601 & credit_scores <= 660, "Poor",
                     ifelse(credit_scores >= 661 & credit_scores <= 720, "Fair",
                     ifelse(credit_scores >= 721 & credit_scores <= 780, "Good",
                     ifelse(credit_scores >= 781 & credit_scores <= 850, "Excellent", NA)))))
```

\
```{r}
category_counts <- table(credit_categories)

category_counts
```